---
title: "Embedding-based Disease Distance Metric"
author: "Joy_Fu"
date: '2022-07-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic setups
```{r Setups, message=FALSE, warning=FALSE, paged.print=FALSE}
rm(list = ls())
pacman::p_load(tidyverse, readr, WGCNA, cluster, fossil)

raw_data_path = "/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/data/"
output_path = "/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/outputs/"
# Source in useful functions
source("/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/code/funcs_used.R")
```

# Part 1. Calculate embedding similarity
This will take a long time, already generated the final outputs in the data folder.
```{r}
source("/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/code/Embed_Dist/1.1_RV_embedding_similarity.R")
# Output file: /Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/data/Embed_Dist/mod/embed_sim.rda
```

# Part 2. Data Cleaning
```{r}
load(file = paste0(raw_data_path, 'Embed_Dist/mod/embed_sim.rda'))
dim(embed_sim)
# We first want to calibrate the RV coefficients to [0,1]
embed_sim_cal = range01(embed_sim)
# Get long format (pairwise distance)
embed_sim_full = makeSymm(embed_sim_cal)
embed_long = reshape2::melt(embed_sim_full)
colnames(embed_long) = c("ICD1", "ICD2", "embed_sim")
embed_long_df = embed_long %>% as.data.frame() %>% 
  mutate(embed_dist = 1 - embed_sim)
save(embed_long_df, file = paste0(raw_data_path, 'Embed_Dist/mod/embed_long_df.rda'))
```

# Part 3. Clustering
## 1. Hierarchical static cut tree
```{r}
embed.dist = as.dist(1 - embed_sim_cal)
consTree = hclust(embed.dist, method = "complete")
# Hierarchical tree only
hclusters = cutree(consTree, k = 22)
table(hclusters)
```

## 2. K-medoids
```{r}
embed.kmedoids = pam(embed.dist, 22) # create k-medoids clustering with 22 clusters
kclusters = embed.kmedoids$cluster
table(kclusters)
```

## 3. Hierarchical dynamic cut trees
```{r}
# WGCNA w/ Dynamic Tree Cut
# Set min module size -- here we set the number to which can result in 22 clusters
minModuleSize = 41
unmergedLabels = cutreeDynamic(dendro = consTree, cutHeight = NULL,
                               distM = 1 - embed_sim_full,
                               minClusterSize = minModuleSize)
unmergedColors = labels2colors(unmergedLabels)
table(unmergedLabels)
```

## 4. Evaluation of clustering results
### 1) Record clustering results
```{r Record the cluster results}
cluster_result = cbind(colnames(embed_sim_cal), hclusters, kclusters, unmergedLabels, unmergedColors)
colnames(cluster_result) = c("ICD", "hcluster_embed", "kcluster_embed", "dtcluster_embed", "dtcolor_embed")
cluster_result_addChap = cluster_result %>% as.data.frame() %>% 
  mutate(chap_clinic = case_when(
    substr(ICD, 1, 1) == "A" | substr(ICD, 1, 1) == "B" ~ "I",
    substr(ICD, 1, 1) == "C" | substr(ICD, 1, 2) %in% c(paste0("D", 0:4)) ~ "II",
    substr(ICD, 1, 2) %in% c(paste0("D", 5:9)) ~ "III", 
    substr(ICD, 1, 1) == "E" ~ "IV",
    substr(ICD, 1, 1) == "F" ~ "V",
    substr(ICD, 1, 1) == "G" ~ "VI",
    substr(ICD, 1, 2) %in% c(paste0("H", 0:5)) ~ "VII",
    substr(ICD, 1, 2) %in% c(paste0("H", 6:9)) ~ "VIII",
    substr(ICD, 1, 1) == "I" ~ "IX",
    substr(ICD, 1, 1) == "J" ~ "X",
    substr(ICD, 1, 1) == "K" ~ "XI",
    substr(ICD, 1, 1) == "L" ~ "XII",
    substr(ICD, 1, 1) == "M" ~ "XIII",
    substr(ICD, 1, 1) == "N" ~ "XIV",
    substr(ICD, 1, 1) == "O" ~ "XV",
    substr(ICD, 1, 1) == "P" ~ "XVI",
    substr(ICD, 1, 1) == "Q" ~ "XVII",
    substr(ICD, 1, 1) == "R" ~ "XVIII",
    substr(ICD, 1, 1) == "S" | substr(ICD, 1, 1) == "T" ~ "XIX",
    substr(ICD, 1, 1) %in% c("V", "W", "X", "Y") ~ "XX",
    substr(ICD, 1, 1) == "Z" ~ "XXI",
    substr(ICD, 1, 1) == "U" ~ "XXII"
  )) %>% 
  mutate(chapter_num = as.numeric(as.roman(chap_clinic))) %>% 
  mutate_at(c("hcluster_embed", "kcluster_embed", "dtcluster_embed"), as.numeric)

# Output results to table
write.table(cluster_result, file = paste0(output_path, "ICD_embed_cluster_results.txt"), 
            sep = "\t", quote = F, row.names = F, col.names = T)
# save(cluster_result, file = paste0(raw_data_path, "Embed_Dist/mod/cluster_result.rda"))
```

### 2) Rand index
```{r Rand index calculation}
# Calculate rand index
rand_hcluster_e = rand.index(cluster_result_addChap$hcluster_embed, cluster_result_addChap$chapter_num)
rand_kmed_e = rand.index(cluster_result_addChap$kcluster_embed, cluster_result_addChap$chapter_num)
rand_dtc_e = rand.index(cluster_result_addChap$dtcluster_embed, cluster_result_addChap$chapter_num)

results_cluster = c(rand_hcluster_e, rand_kmed_e, rand_dtc_e)
results_cluster
```

# Appendix -- Get joint_CONCEPT.rda (just need to run once)
```{r joint.CONCEPT.rda}
# Read in raw mapping data downloaded from Athena
CONCEPT = read_delim(paste0(raw_data_path, "Raw/Embedding/athena.download.1209/CONCEPT.csv"), delim = "\t", escape_double = FALSE, trim_ws = TRUE)
CONCEPT_RELATIONSHIP = read_delim(paste0(raw_data_path, "Raw/Embedding/athena.download.1209/CONCEPT_RELATIONSHIP.csv"), delim = "\t", escape_double = FALSE, trim_ws = TRUE)

CONCEPT_ICD = CONCEPT %>% 
  filter(vocabulary_id == "ICD10CM" | vocabulary_id == "ICD10") %>% 
  filter(domain_id == "Condition") %>% 
  select(concept_id, concept_name, concept_code) %>% unique()
CONCEPT_SNOMED = CONCEPT %>% 
  filter(vocabulary_id == "SNOMED" & domain_id == "Condition") %>% 
  select(concept_id, concept_code)
CONCEPT_RELATIONSHIP_short = CONCEPT_RELATIONSHIP %>% 
  filter(relationship_id == "Maps to") %>% 
  filter(concept_id_1 != concept_id_2) %>% 
  select(concept_id_1, concept_id_2)
# Join together
CONCEPT_RELATIONSHIP_full = CONCEPT_RELATIONSHIP_short %>% 
  left_join(CONCEPT_SNOMED, by = c("concept_id_2" = "concept_id")) %>% 
  left_join(CONCEPT_ICD, by = c("concept_id_1" = "concept_id")) %>% 
  rename(ICD_10 = concept_code.y,
         SNOMED_id = concept_code.x) %>% drop_na() %>% select(-c(concept_id_1, concept_id_2))
joint_CONCEPT = CONCEPT_RELATIONSHIP_full %>% select(-concept_name) %>% 
  unique() %>% arrange(ICD_10)

save(joint_CONCEPT, file = paste0(raw_data_path, 'Embedding.Sim/mod/joint_CONCEPT.rda'))
```


