---
title: "Comorbidity-based Disease Distance Metric"
author: "Joy_Fu"
date: '2022-07-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic setups
```{r Setups, message=FALSE, warning=FALSE, paged.print=FALSE}
rm(list = ls())
pacman::p_load(tidyverse)

raw_data_path = "/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/data/"
output_path = "/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/outputs/"
# Source in useful functions
source("/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/code/funcs_used.R")
```

# Calculate comorbidity between disease pairs (EHR data on server)
File path: /opt/genomics/IPHinvestigators/joyfu/disease_similarity/code/comorbidity/code/
* 3.1_hz_prepare.R (Clean visit.tsv data, prepare dataset for Cox model analyses)
* 3.2_full_tb_hz_prepare.R (Each diagnosis code as a separate file, record time)
* 3.3_file_join.R (Join all diagnosis files together)
* 3.4_run_hz_sep.R (Get comorbid data, also as a filtering step for HR calculation)

File path: /opt/genomics/IPHinvestigators/joyfu/disease_similarity/code/comorbidity/code/run_sep/
* Separate_DF.ipynb (separate running files to reduce memory)
* xx_run_hz_sep.R (Run HZ model separately)

File path: /opt/genomics/IPHinvestigators/joyfu/disease_similarity/code/comorbidity/data/new/output/
* Combine_results.ipynb (Combine separate results)

Output file: 
* Comorbidity in UCLA EHR: /Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/data/Comorb_Dist/comorb_df_long.rda



