---
title: "Genetic Correlation-based Disease Distance Metric"
author: "Joy_Fu"
date: '2022-07-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Setups}
rm(list = ls())
pacman::p_load(tidyverse, readr, readxl, gtsummary)

raw_data_path = "/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/data/"
output_path = "/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/outputs/"
# Source in useful functions
source("/Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/code/funcs_used.R")
```

# Part 1. Calculate genetic correlation w/ UKBB GWASs 
File path: /Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/code/Genetic_Dist/
* 2.1_loop_GC.sh (Pairwise genetic correlation using LDSC tool)
* script_GC_computation.sh (no need to run, incorporated in the 2.1_loop_GC.sh)
* 2.2_script_summarize_GC.sh (summarizing genetic correlation results)
Output file: /Users/Mingzhou/Desktop/Projects/Disease.Similarity/GitHub/data/Genetic_Dist/mod/Genetic_Correlation_of_UKB_diseases.txt

# Part 2. Cleaning LDSC results
```{r message=FALSE, warning=FALSE}
# Load in LDSC results
gen_result = read.delim(file = paste0(raw_data_path, "Genetic_Dist/mod/Genetic_Correlation_of_UKB_diseases.txt"), header = TRUE, sep = " ")
# Restrict to 3-digit ICD only, remove NAs
gen_3digit = gen_result %>% filter(!is.na(rg)) %>% 
  filter(nchar(p1) == 3 & nchar(p2) == 3) %>% select(p1, p2, rg, se, z, p) %>% unique()
gen_3digit_add = gen_3digit %>% dplyr::rename(p1 = p2, p2 = p1) 
gen_3digit_full = gen_3digit %>% full_join(gen_3digit_add)
# N = 182,754
# Also load in the heritability files (from Neals Lab)
h2_ukbb_raw = read.table(file = paste0(raw_data_path, "Genetic_Dist/raw/UKBB_h2info_download.tsv"), sep = '\t', header = TRUE)
```

## 1. Check associations w/ Neale's Lab's results
```{r}
# Load in data
neale_cor_raw = read_csv(file = paste0(raw_data_path, "Genetic_Dist/raw/NealeLab-gencor-2022-01-31.csv"))
neale_cor = neale_cor_raw %>% select(ID1, ID2, h2, `h2 SE`, rg, Z, p) %>% 
  dplyr::rename(h2_neale = h2, h2_se_neale = `h2 SE`, rg_neale = rg, Z_neale = Z, p_neale = p)
# Combine data
gen_check = gen_3digit_full %>% 
  left_join(neale_cor, by = c("p1" = "ID1", "p2" = "ID2"))
# Check correlations
cor.test(gen_check$rg, gen_check$rg_neale)
cor.test(gen_check$rg, gen_check$rg_neale, method = "spearman")
```

## 2. Filter by h2 (h2_liability, not observed)
```{r}
UKBB_raw = read_excel(paste0(raw_data_path, "Genetic_Dist/raw/UKBB_Imputed_v3_20180731.xlsx"), sheet = "Manifest 201807")
UKBB_gwas = UKBB_raw %>% 
  filter(substr(`Phenotype Code`, 1, 1) %in% LETTERS) %>% 
  filter(Sex == "both_sexes") %>% 
  select(-c(`UK Biobank Data Showcase Link`, `md5s`)) %>% 
  filter(`Phenotype Code` != "N/A" & `Phenotype Description` != "NA") %>% 
  mutate(`Phenotype Description` = tolower(`Phenotype Description`)) %>% 
  select(`Phenotype Code`, `wget command`)
colnames(UKBB_gwas) = c("phenotype", "wget_command")
# Load in the GWAS lists we extracted
wget_lst = read.csv2(file = paste0(raw_data_path, "Genetic_Dist/raw/wget_lst.csv"), header = F)
colnames(wget_lst) = "wget_command"
# Join these two
UKBB_gwas_filter = UKBB_gwas %>% right_join(wget_lst)
# Preprocess of h2
h2_ukbb_icd = h2_ukbb_raw %>% 
  filter(substr(phenotype, 1, 1) %in% LETTERS) 
# combine two
h2_combine = h2_ukbb_icd %>% right_join(UKBB_gwas_filter) %>% select(-wget_command) %>% 
  filter(!is.na(h2_liability)) 
# Filter by h2 reliability
h2_valid = h2_combine %>% 
  filter(confidence %in% c("medium", "high")) %>% 
  select(phenotype, h2_liability, h2_liability_se, confidence)
dis_lst = h2_valid %>% pull(phenotype)
dis_lst = c(dis_lst, "M20", "M54") %>% unique()
# N = 31
gen_sim_check = gen_3digit_full %>% 
  filter(p1 %in% dis_lst & p2 %in% dis_lst) %>% 
  mutate(p_adj_fdr = p.adjust(p, method = "fdr")) %>% 
  left_join(h2_valid, by = c("p1" = "phenotype")) %>% 
  dplyr::rename(H2_d1 = h2_liability, H2_d1_se = h2_liability_se, d1_confidence = confidence) %>% 
  left_join(h2_valid, by = c("p2" = "phenotype")) %>% 
  dplyr::rename(H2_d2 = h2_liability, H2_d2_se = h2_liability_se, d2_confidence = confidence)
# N = 756 pairs (without filtering by FDR p-value)
```

# Part 3. Cleaning mining genetic correlation results
## 1. Load in corr and h2 info, map together
```{r}
# Load in data
mine_corr = read_excel(paste0(raw_data_path, "Genetic_Dist/raw/mine_corr.xlsx"))
mine_corr_short = mine_corr %>% 
  dplyr::rename(corr = Correlation_Value, corr_SD = Correlation_SE, published = `Published(1)/Predicted(0)`,
         Correlation_type_in_reference = Correlation_Type, Data_type_in_reference = Data_Type, 
         Math_model_in_reference = Math_Model, Country_of_cohort = Country_of_Cohort,
         Gender_of_cohort = Gender_of_Cohort) %>% 
  filter(published == 1 & Gender_of_cohort == "Both" & Correlation_type_in_reference == "Genetic correlation") %>% 
  filter(Data_type_in_reference != "Polygenic risk score") %>% 
  select(Disease1, Disease2, corr, corr_SD, Data_type_in_reference, Math_model_in_reference, PMID)
# N = 1055
# Also add h2 information
mine_h2 = read_excel(paste0(raw_data_path, "Genetic_Dist/raw/mine_h2.xlsx"))
mine_h2_short = mine_h2 %>% 
  dplyr::rename(H2 = `Heritability Value`, H2_se = `Heritability SE`, published = `Published(1)/Predicted(0)`,
         Data_type = `Data Type`, Math_model_in_reference = `Math Model`,
         Country_of_cohort = `Country of Cohort`, Gender_of_cohort = `Gender of Cohort`) %>% 
  filter(published == 1 & Gender_of_cohort == "Both") %>% 
  filter(Data_type != "Polygenic risk score") %>% 
  select(Disease, H2, H2_se, Data_type, Math_model_in_reference, PMID)
# N = 975
# Join those info together by PMID
mine_corr_addh2 = mine_corr_short %>% 
  left_join(mine_h2_short, by = c("Disease1" = "Disease", "PMID" = "PMID",
                                  "Data_type_in_reference" = "Data_type", 
                                  "Math_model_in_reference" = "Math_model_in_reference")) %>% 
  dplyr::rename(H2_D1 = H2, H2_se_D1 = H2_se) %>% 
  left_join(mine_h2_short, by = c("Disease2" = "Disease", "PMID" = "PMID",
                                  "Data_type_in_reference" = "Data_type", 
                                  "Math_model_in_reference" = "Math_model_in_reference")) %>% 
  dplyr::rename(H2_D2 = H2, H2_se_D2 = H2_se)
```

## 2. Add ICD-10 code to diseases
```{r}
# Read in an outside file including all disease-ICD mapping
disease_tomap = read.table(file = paste0(raw_data_path, 'Genetic_Dist/raw/disease_tomap.txt'), header = F, sep = "\t")
names(disease_tomap) = c("Disease", "ICD.10")
# Add mapping
mine_corr_addh2_ICD = mine_corr_addh2 %>% 
  left_join(disease_tomap, by = c("Disease1" = "Disease")) %>% 
  mutate(ICD_1 = substr(ICD.10, 1, 3)) %>% 
  left_join(disease_tomap, by = c("Disease2" = "Disease")) %>% 
  mutate(ICD_2 = substr(ICD.10.y, 1, 3)) %>% 
  filter(corr_SD != 0) %>% 
  select(ICD_1, ICD_2, corr, corr_SD, Data_type_in_reference, Math_model_in_reference, H2_D1, H2_se_D1, H2_D2, H2_se_D2) 
# Pair check
pair_check = mine_corr_addh2_ICD %>% group_by(ICD_1, ICD_2, Data_type_in_reference) %>% 
  dplyr::summarise(n = n(), min_sd = min(corr_SD)) 
mine_corr_addh2_ICD_check = mine_corr_addh2_ICD %>% left_join(pair_check) %>% 
  mutate(both_h2 = case_when(
    !is.na(H2_D1) & !is.na(H2_D2) ~ 1,
    TRUE ~ 0
  )) 
```

## 3. Filtering results (keep one for each pair)
```{r}
mine_corr_keep = mine_corr_addh2_ICD_check %>% 
  mutate(if_keep = case_when(
    n == 1 ~ 1,
    n > 1 & both_h2 & (corr_SD == min_sd) ~ 2,
    n > 1 & (corr_SD == min_sd) ~ 3,
    TRUE ~ 0
  )) %>% filter(if_keep != 0) %>% 
  group_by(ICD_1, ICD_2, Data_type_in_reference) %>% slice(1)
# N = 689 pairs
mine_corr_keep2 = mine_corr_keep %>% 
  dplyr::rename(ICD_2 = ICD_1, ICD_1 = ICD_2, H2_D2 = H2_D1, H2_D1 = H2_D2, 
                H2_se_D2 = H2_se_D1, H2_se_D1 = H2_se_D2)
mine_corr_full = mine_corr_keep %>% full_join(mine_corr_keep2)
# N = 1374 pairs
```

# Part 4. Join LDSC and mined corr, filter valid ones
```{r}
gen_corr_full = gen_sim_check %>% 
  select(p1, p2, rg, se, z, p, p_adj_fdr, H2_d1, H2_d2, d1_confidence, d2_confidence) %>% 
  full_join(mine_corr_full, by = c("p1" = "ICD_1", "p2" = "ICD_2")) %>% 
  mutate(Study_type = case_when(
    Data_type_in_reference %in% c("Family study", "Family study using EHR", "Twin study") ~ "Family/Twin-study",
    TRUE ~ "SNP-based"
  )) %>% filter(!is.na(p1) & !is.na(p2)) %>% rowwise() %>% 
  mutate(gen_cor_sd = case_when(
    !is.na(rg) & !is.na(corr) ~ min(se, corr_SD, na.rm = T),
    !is.na(rg) ~ se,
    !is.na(corr) ~ corr_SD
  )) %>% 
  mutate(gen_cor = case_when(
    gen_cor_sd == se ~ rg,
    gen_cor_sd == corr_SD ~ corr
  )) %>% filter(p_adj_fdr < 0.1 | !is.na(corr)) %>% 
  mutate(h2_D1 = case_when(
    gen_cor_sd == se ~ H2_d1,
    gen_cor_sd == corr_SD ~ H2_D1
  )) %>% 
  mutate(h2_D2 = case_when(
    gen_cor_sd == se ~ H2_d2,
    gen_cor_sd == corr_SD ~ H2_D2
  )) %>% 
  select(p1, p2, gen_cor, gen_cor_sd, Study_type, h2_D1, d1_confidence, h2_D2, d2_confidence) %>% 
  dplyr::rename(ICD1 = p1, ICD2 = p2) %>% unique() %>% filter(ICD1 != ICD2)
gen_corr_full = as.data.frame(gen_corr_full)
# N = 1806 (total pairs)
gen_corr_single = gen_corr_full %>% rowwise() %>% 
  mutate(id = paste(sort(c(ICD1, ICD2)), collapse = "-")) %>% ungroup() 
gen_corr_final = gen_corr_single %>% 
  group_by(id, gen_cor, gen_cor_sd, Study_type) %>% slice(1) %>% ungroup() %>% select(-id)
# N = 903 (single pairs)
# Save results
save(gen_corr_full, file = paste0(raw_data_path, "Genetic_Dist/mod/gen_corr_full.rda"))
save(gen_corr_final, file = paste0(raw_data_path, "Genetic_Dist/mod/gen_corr_final.rda"))
```

## 1. Check genetic results
###  Distributions
```{r}
summary(gen_corr_full$gen_cor)
```

### Check distributions by study type
```{r}
gene_compare_short = gen_corr_full %>% as.data.frame() %>%
  select(gen_cor, gen_cor_sd, Study_type) %>% unique()
study_compare = tbl_summary(
  gene_compare_short,
  by = Study_type
) %>% add_n() %>% add_p()
study_compare
```